---
title: "Sample Selection for Proteomics"
output:
    flexdashboard::flex_dashboard:
        navbar:
            - { title: "Home", 
                href: "../../site/_site/hdl.html", 
                align: left }
        source_code: embed
runtime: shiny 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning =F, error = F, message=F)
```

```{r, packages}
pkgs = c('plyr', 'dplyr','stringr','reshape2','tibble', 'plotly', 'DT',
         'ggplot2','ggsci','knitr')
for(pkg in pkgs){
    library(pkg, quietly=TRUE, verbose=FALSE, warn.conflicts=FALSE, 
            character.only=TRUE)
}
```

```{r}
load("../../data/diet.Rdata")
load("../Rdata/fct_precalc.Rdata")
```

```{r}
df = data.frame(
    chol_efflux = fct_data$`Chol Efflux (ApoB Depleted Serum 1%)`,
    Subj = pdata$Subj,
    Day = pdata$Day,
    TX = pdata$TX
) %>%
    filter(TX=="egg") %>%
    dcast(Subj + TX ~ Day, value.var="chol_efflux") %>%
    mutate(change = Post-Pre)
```

Inputs {.sidebar}
------------------------------

```{r}
inputPanel(
    numericInput("n","QC Sample Size (n)", value = 2),
    numericInput("sigma", "QC Standard Deviation", value = 2, step=1),
    numericInput("alpha", "alpha", value = 0.05),
    numericInput("seed", "Random seed", value=12321),
    selectInput("var", "Select a Variable",
                choices = names(clinical_data), 
                selected="Cholesterol total (mg/dL)")
)
```

Column
-------------------------------

### The Density Distribution of Chol Efflux Measurement {data-height=500}

```{r}
dnorm_shade = function(cutoff, x, mean, sd){
    y = dnorm(x, mean, sd)
    y[x<cutoff] = NA
    return(y)
}
cutoff = reactive({
    qnorm((1-input$alpha), mean=0, sd=input$sigma/sqrt(input$n))
})
renderPlot({
    sd = input$sigma/sqrt(input$n)
    ggplot(data.frame(mu=c(-3*input$sigma, 3*input$sigma)), aes(x=mu)) +
        stat_function(fun=dnorm, geom="line", color = "darkblue",size=1,
                      args=list(mean=0, sd=sd))+
        stat_function(fun=dnorm_shade, geom="area", fill="red", alpha=0.25,
                      args=list(cutoff=cutoff(), mean=0, 
                                sd=sd)) +
        geom_vline(xintercept = cutoff(), linetype="dashed", size=1, 
                   color="darkred") +
        theme_bw()
})
```

### {data-height=330}

The difference needs to be bigger than **`r reactive({cutoff()})`** so we are able to reject the null hypothesis that the two samples are the same. Subjects with cholesterol efflux increased more than **`r reactive({cutoff()})`** are considered as responder.

#### Responder:

```{r}
responder = reactive({as.character(df$Subj[df$change > cutoff()])})
reactive({cat(responder())})
```

#### Non-responder:

```{r}
non_responder = reactive({as.character(df$Subj[df$change <= cutoff()])})
reactive(cat(non_responder()))
```

### Samples Random Selection {data-height=170}

```{r, include=F}
responder_selected = reactive({
    set.seed(input$seed)
    sample(responder(),5)
})
non_responder_selected = reactive({
    set.seed(input$seed)
    sample(non_responder(),5)
})
```

```{r}
renderTable({
    data.frame(responder_selected(), non_responder_selected()) %>% t %>% as.data.frame
}, rownames=T, colnames=F)
```

Column
--------------------

### Chol Efflux (all subj) {data-height=550}

```{r}
renderPlotly({
    mycolors = colorRampPalette(pal_jama()(7))(20)
    df2 = data.frame(
        var = fct_data$`Chol Efflux (ApoB Depleted Serum 1%)`,
        Subj = pdata$Subj,
        TX = pdata$TX,
        Day = pdata$Day
    )
    p = ggplot(df2, aes(Day, var)) +
        geom_boxplot() +
        geom_point(aes(color = Subj), size=2) +
        geom_line(aes(group=Subj, color=Subj)) +
        facet_grid(.~TX) +
        scale_color_manual(values=mycolors) +
        labs(y="") +
        theme_bw()
    ggplotly(p)
})
```

### Selected Subjs {data-height=450}


```{r}
renderPlotly({
    mycolors = colorRampPalette(pal_jama()(7))(20)
    df2 = data.frame(
        var = clinical_data[,input$var],
        Subj = pdata$Subj,
        TX = pdata$TX,
        Day = pdata$Day
    ) %>%
        mutate(
            Label = ifelse(
                Subj %in% responder_selected(), "Responder",
                ifelse(
                    Subj %in% non_responder_selected(), "Non-responder", NA
                ))) %>%
        filter(!is.na(Label))
    p = ggplot(df2, aes(Day, var)) +
        geom_boxplot() +
        geom_point(aes(color = Subj)) +
        geom_line(aes(group = Subj, color = Subj)) +
        facet_grid(Label~TX) +
        scale_color_manual(values=mycolors) +
        labs(y="", title=input$var) +
        theme_bw()
    ggplotly(p)
})
```


