---
title: "Egg Study DIA Exploratory"
author: "Chenghao Zhu"
date: "9/28/2018"
output: html_document
---

<style>
p {
  background-color: #eee;
  border: 1px solid #999;
  display: block;
  padding: 20px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
```

```{r}
pkgs = c("dplyr", "reshape2", "stringr", "tibble", "Metabase", "readr", "DT", 
         "ggplot2", "plotly")
for (pkg in pkgs) {
    library(pkg, character.only = TRUE, verbose = FALSE, quietly = TRUE, warn.conflicts = FALSE)
}
```

```{r}
file = "../../../raw_data/proteome/Samples Report of 20180710-Lisa-fullsetDIA.csv"
df = read_csv(file, skip = 13, col_names = FALSE, na = "Missing Value")
df = df[-nrow(df),]
names = read_csv(file, skip = 12, col_names = FALSE, n_max = 1) %>%
    as.character()

fdata = df[,1:9]
colnames(fdata) = names[1:9]

count = df[10:29]
colnames(count) = paste0(
    "Egg", gsub(".+Lisa_([0-9]{3}[A-D]{1})_DIA_wide.raw", "\\1", names[10:29])
)

log_int = df[50:69]
colnames(log_int) = paste0(
    "Egg", gsub(".+Lisa_([0-9]{3}[A-D]{1})_DIA_wide.raw", "\\1", names[50:69])
)

pdata = data.frame(
    row.names = colnames(count),
    Subject = gsub("^Egg(\\d{3})[A-D]{1}$", "\\1", colnames(count)),
    Timepoint = ifelse(
        gsub("^Egg\\d{3}([A-D]{1})", "\\1", colnames(count)) %in% c("A", "C"),
        "Pre", "Post"
    ),
    Responding = c(rep("Non-responder", 10), rep("Responder", 10))
)
pdata$Subject = factor(pdata$Subject)
pdata$Timepoint = factor(pdata$Timepoint, levels = c("Pre", "Post"))
pdata$Responding = factor(pdata$Responding, levels = c("Non-responder", "Responder"))

count = as.matrix(count)
rownames(count) = str_c("Feature", str_pad(1:nrow(count), width=3, pad="0"))

log_int = as.matrix(log_int)
rownames(log_int) = rownames(count)

fdata = fdata[,-1]
fdata = as.data.frame(fdata)
rownames(fdata) = rownames(count)
fdata = fdata[!grepl("^DECOY", fdata$`Protein Name`),]
count = count[rownames(fdata),]
log_int = log_int[rownames(fdata),]
fdata$Species = gsub(".+OS=(.+?) [A-Z]{2}=.+", "\\1", fdata$`Protein Name`)
fdata$Protein = gsub(
    "^[st][pr]|\\S+? (.+?) OS=.+", "\\1",
    gsub("^Group of ", "", fdata$`Protein Name`)
)

count = ProteomicsSet(
    conc_table = conc_table(count),
    sample_table = sample_table(pdata),
    feature_data = feature_data(fdata)
)

log_int = ProteomicsSet(
    conc_table = conc_table(log_int),
    sample_table = sample_table(pdata),
    feature_data = feature_data(fdata)
)

featureNames(count) = count$feature_data$`Accession Number`
featureNames(log_int) = log_int$feature_data$`Accession Number`
```

## Part I Data Cleaning

The csv file "Samples Report of 20180710-Lisa-fullsetDIA.csv" was used. The original data has 224 variables. The variable "DECOY_sp|P04114|APOB_HUMAN" was removed. 

The mean of count for responders and non-responders before and after treatment.

```{r}
count$conc_table %>%
    t %>% as.tibble %>%
    mutate(
        Subject = count$sample_table$Subject,
        Timepoint = count$sample_table$Timepoint,
        Responding = count$sample_table$Responding
    ) %>%
    melt(id.var = c("Subject", "Timepoint", "Responding")) %>%
    group_by(Timepoint, Responding, variable) %>%
    summarize(value = mean(value)) %>%
    dcast(variable ~ Timepoint + Responding) %>%
    mutate(Protein = count$feature_data[variable, "Protein"]) %>%
    column_to_rownames("variable") %>%
    datatable()
```

The mean of intensity for responders and non-responders before and after treatment.

```{r}
log_int$conc_table %>%
    t %>% as.tibble %>%
    mutate(
        Subject = log_int$sample_table$Subject,
        Timepoint = log_int$sample_table$Timepoint,
        Responding = log_int$sample_table$Responding
    ) %>%
    melt(id.var = c("Subject", "Timepoint", "Responding")) %>%
    group_by(Timepoint, Responding, variable) %>%
    summarize(value = mean(value, na.rm = TRUE)) %>%
    dcast(variable ~ Timepoint + Responding) %>%
    mutate(Protein = log_int$feature_data[variable, "Protein"]) %>%
    column_to_rownames("variable") %>%
    datatable()
```

Below is a barplot of number of proteins with missing values. There are 9 variables with 1 missing values, 5 with 2 missing values, and there are 81 proteins that none of the samples has a value.

```{r}
plot_hist_NA(log_int)
```

Proteins with morer than 5 missing values were removed. For the rest proteins, missing values were filled up using 1/2 of the lowest value.

```{r}
log_int = subset_features(
    log_int, apply(log_int$conc_table, 1, function(x) sum(is.na(x)) <= 5))
count = subset_features(count, featureNames(log_int))
log_int = transform_by_feature(
    log_int, function(x){
        ifelse(is.na(x), min(x, na.rm = TRUE)/2, x)
    })
```

Below is the barplot of the sum of each sample's count. This shows that the total amount of peptides are relatively equivalent between samples.

```{r}
par(mfrow = c(1,2))
barplot(colSums(count$conc_table), main = "Count")
barplot(colSums(log_int$conc_table), main = "Log Intensity")
```

## Part 2, statistic analysis

```{r}
design = model.matrix(data = as(count$sample_table, "data.frame"), 
                      ~Responding * Timepoint + Subject + 1)
# de = mSet_deseq(count, design, result = "RespondingResponder")
lm = mSet_limma(log_int, design, function(x){x}, coef = 13, p.value = 13) %>%
    rownames_to_column("Access Number") %>%
    mutate(Protein = log_int$feature_data$Protein)
```


Linear model was performed on the log-transformed intensity data, using the formular below:

~ Responding + Timepoint + Responding * Timepoint + Subject + 1

The Responding * Timepoint coefficient was tested.

Below is the distribution of the p values

```{r}
ggplot(lm) +
    geom_histogram(aes(pvalue), binwidth = 0.05, color = "white") +
    theme_bw()
```

Vocano plot. Only one protein, Anthrex toxin receptor 1, has a p value < 0.05. Not a typical HDL protein.

```{r}
(ggplot(lm) +
    geom_point(aes(logFC, -log(pvalue), `Access Number`=`Access Number`, Protein=Protein), 
               color = "white", 
               fill = "gray17", shape=21, size=3) +
    theme_bw()) %>%
    ggplotly
```

Below is the statistic result for each protein. 

```{r}
lm %>%
    arrange(pvalue) %>%
    select(`Access Number`, Protein, baseMean, logFC, stat, pvalue, padj) %>%
    sapply(function(x){
        if(is.character(x)) return(x)
        round(x, 3)
    }) %>%
    datatable()
```

Boxplot of Anthrax toxin receptor 1

```{r}
(plot_boxplot(log_int, feature = "sp|Q9H6X2|ANTR1_HUMAN",x = "Timepoint", cols = "Responding", line = "Subject", color = "Subject") +
    labs(title = "Anthrax toxin receptor 1")) %>%
    plotly::ggplotly()
```